Combinatorial Proofs and Decomposition Theorems for First-order Logic

### 2021-01-13 main.tex
sec.VII and VIII

\usepackage{xifthen}
\usepackage[utf8]{inputenc}
\usepackage{dashbox}
\usepackage{ebproof}
\usepackage{amsmath}
\usepackage{ebproof}
\usepackage{virginialake}

@@

\newcommand{\tuple}[1]{\langle#1\rangle}
\newcommand{\pair}[1]{(#1)}
\newcommand{\set}[1]{\{#1\}}
\newcommand{\sqn}[1]{\vdash#1}

@@

\newcommand{\rectif}[1]{\widehat{#1}}
\newcommand{\graphof}[1]{\llbracket#1\rrbracket}
\newcommand{\compl}[1]{#1^\complement}
%\newcommand{\sublist}[1]{\set{#1}}
%\newcommand{\subst}[2]{#1\mathord{\scriptstyle\mapsto} #2}
\newcommand{\sublist}[1]{[#1]}
\newcommand{\subst}[2]{#1/#2}
\newcommand{\ssubst}[2]{\sublist{\subst{#1}{#2}}}

@@

\section{Introduction}

@@

\begin{lemma}
  Let $A$ and $B$ be formulas. Then
  \begin{equation}
    x
  \end{equation}
\end{lemma}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{A First Decomposition Theorem}

$$
\vlderivation{
  \vlde{\set{\wD,\cD}}{}{C}{
    \vlde{\FOMLS}{}{C'}{
      \vlhy{\ttt}}}}
$$

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Unification Nets}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Translation between Unification Nets and $\FOMLL$}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Translation between Unification Nets and Fonets}

>>>>

\usepackage{xifthen}
\usepackage[utf8]{inputenc}
\usepackage{dashbox}
\usepackage{ebproof}
\usepackage{amsmath}
\usepackage{ebproof}
\usepackage{virginialake}
\vlstemheight=12pt

@@

\newcommand{\tuple}[1]{\langle#1\rangle}
\newcommand{\pair}[1]{(#1)}
\newcommand{\set}[1]{\{#1\}}
\newcommand{\sqn}[1]{\vdash#1}
\newcommand{\sqns}[1]{\vdash#1\phantom{\vdash}}

@@

\newcommand{\rectif}[1]{\widehat{#1}}
\newcommand{\graphof}[1]{\llbracket#1\rrbracket}
\newcommand{\compl}[1]{#1^\complement}
%\newcommand{\sublist}[1]{\set{#1}}
%\newcommand{\subst}[2]{#1\mathord{\scriptstyle\mapsto} #2}
\newcommand{\sublist}[1]{[#1]}
\newcommand{\subst}[2]{#1/#2}
\newcommand{\ssubst}[2]{\sublist{\subst{#1}{#2}}}
\newcommand{\substof}[1]{\sigma(#1)}
\newcommand{\linkingof}[1]{\sim_{#1}}

@@

\todo{Examples,examples,examples}
\section{Introduction}

@@

\begin{lemma}
  Let $A$ and $B$ be formulas. Then
  \begin{equation*}
    \vlderivation{
      \vlde{\set{\wlD,\wrD,\cD,\fequ}}{}{B}{
        \vlhy{A}}}
    \qquad
    \iff
    \qquad
    \vlderivation{
      \vlde{\set{\wlD,\wrD,\acD,\me,\mfaD,\mexD,\fequ}}{}{B}{
        \vlhy{A}}}
  \end{equation*}
\end{lemma}

\begin{proof}
  This follows immediately from Lemmas~\ref{lem:ac} and~\ref{lem:me}.
\end{proof}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Rule Permutations}

%All statements in this section are proved by rule permutation arguments.
\begin{thm}\label{thm:LK1-decompose}
  Let $\Gamma$ be a sequent. If\/ $\sqn\Gamma$ is provable in $\FOLK$ (as depicted on the left below) then there is a sequent
  $\Gamma'$ such that there is a derivation as shown on the right below:\vadjust{\vskip-2ex}
  \begin{equation*}
    \vlderivation{
        \vltrl{}{\FOLK}{\Deri}{\sqns{\Gamma}}{
          \vlhy{\quad}}{
          \vlhy{}}{
          \vlhy{}}}
    \qquad
    \Longrightarrow
    \qquad
    \vlderivation{
      \vlde{\set{\wlD,\wrD,\cD}}{\Deri_2}{\sqns\Gamma}{
        \vltrl{}{\FOMLL}{\Deri_1}{\sqns{\Gamma'}}{
          \vlhy{\quad}}{
          \vlhy{}}{
          \vlhy{}}}}
  \end{equation*}
\end{thm}

\begin{proof}
  Note that the instances of $\wlD,\wrD,\cD$ in $\Deri_2$ are deep, but inside sequent contexts.
  \todo{}
\end{proof}

\begin{lemma}\label{lem:MLS1-decomposition}
  For every derivation
  $\vlderivation{
      \vlde{\FOMLS}{}{A}{
        \vlhy{\ttt}}}$
  there are formulas $A'$ and $A''$ such that 
  \begin{equation*}
    \vlderivation{
      \vlde{\set{\exists}}{}{A}{
        \vlde{\set{\sw,\mix,\fequ}}{}{A'}{
          \vlde{\set{\forall,\aiD,\tttD}}{}{A''}{
            \vlhy{\ttt}}}}}
  \end{equation*}
\end{lemma}

\begin{lemma}\label{lem:cw-decomposition}
  For every derivation
  $\vlderivation{
      \vlde{\set{\wlD,\wrD,\acD,\me,\mfaD,\mexD,\fequ}}{}{B}{
        \vlhy{A}}}$
  there are formulas $A'$ and $B'$ such that 
  \begin{equation*}
    \vlderivation{
      \vlde{\set{\wlD,\wrD}}{}{B}{
        \vlde{\set{\acD}}{}{B'}{
          \vlde{\set{\me,\mfaD,\mexD,\fequ}}{}{A'}{
            \vlhy{A}}}}}
  \end{equation*}
\end{lemma}

We can now complete the proof of Theorems~\ref{thm:decomposition} and~\ref{thm:decompositionA}.

\begin{proof}[Proof of Theorem~\ref{thm:decompositionA}]
  Assume we have a proof of $A$ in $\FOKS$. By
  Theorem~\ref{thm:LK1-KS1} we have a proof of $\sqn A$ in $\FOLK$ to
  which we can apply Theorem~\ref{thm:LK1-decompose}. Finally, we
  apply Lemma~\ref{lem:MLL1->MLS1} to get the desired shape.
\end{proof}

\begin{proof}[Proof of Theorem~\ref{thm:decomposition}]
  Assume we have a proof of $A$ in $\FOKS$. We first apply
  Theorem~\ref{thm:decompositionA}, and then
  Lemma~\ref{lem:MLS1-decomposition} to the upper half and
  Lemma~\ref{lem:cw-decomposition} to the lower half.
\end{proof}

@@

\todo{}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Unification Nets}

\todo{}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Translation between Unification Nets and $\FOMLL$}

\todo{}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Translation between Unification Nets and Fonets}

\todo{}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Translation between $\FOMLS$ and Fonets}

\todo{define rectification $\rectif\Deri$ of a derivation  $~\vlder{\FOMLS}{\Deri}{A}{\ttt}$}
\todo{define substitution of a derivation $\substof{\Deri}$ and linking of a derivation $\linkingof{\Deri}$ }

\begin{thm}
  Let the derivation $\vlder{\FOMLS}{\Deri}{A}{\ttt}$ be given. Then
  the graph $\graphof{\rectif A}$ forms a fonet whose linking is
  $\linkingof{\rectif\Deri}$ and whose dualizer is
  $\substof{\rectif\Deri}$.
\end{thm}

\begin{thm}
  Let $\gC$ be a fonet with linking $\linkingof\gC$ and dualizer
  $\delta$, and let $A$ be a formula with $\graphof A=\gC$.  Then
  there is a derivation $\vlder{\FOMLS}{\Deri}{A}{\ttt}$ such that
  $\linkingof\gC=\linkingof{\Deri}$ and $\delta=\substof{\Deri}$.
\end{thm}
